Mainstreaming Race Science

### 2023-09-05 paper/paper.Rmd
adding in cites for other professional orgs rejecting race as biology
Added in additional citations to professional anthropology and genetics organizations rejecting "race" as biologically meaningful.

Instead, our results suggest that psychology has been the dominant discipline in which the race science echo chamber has been maintained. This may be the result of racial essentialization being maintained more easily in a discipline like psychology than in other disciplines that contribute to behavior genetics, like genetics and anthropology. *[cite?]* The typical practitioners in psychology do not have the degree of formal education about, and therefore knowledge of, relevant underlying genetic and anthropological research that has led many professional organizations in genetics and anthropology to formally reject the claim of race as meaningful biological concept \cite{anthropology1996aapa, american2018ashg}, undermining the core assumption of race science. As described by Winston, however, "[s]ince the 1990s, scientific racism has evolved in ways that sidestep or blunt the charge of essentialist categorization" \cite[p. 427]{WinstonWhyMainstreamResearch2020}, such as shifting the rhetoric away from race to concepts such as "national IQ" that are deployed to preserve an essentialist categorization of different human populations. Professional psychological organizations routinely publish statements rejecting racism or acknowledging psychology's complicity in systemic racism \cite{american2021apology, aps2021statement}, but stop short of rejecting race as a meaningful biological categorization of people, allowing for race science research to be smuggled into psychology and giving the appearance of continued scientific legitimacy.

>>>>

Instead, our results suggest that psychology has been the dominant discipline in which the race science echo chamber has been maintained. This may be the result of racial essentialization being maintained more easily in a discipline like psychology than in other disciplines that contribute to behavior genetics, like genetics and anthropology. *[cite?]* The typical practitioners in psychology do not have the degree of formal education about, and therefore knowledge of, relevant underlying genetic and anthropological research that has led many professional organizations in genetics and anthropology to formally reject the claim of race as a biologically meaningful concept \cite{anthropology1996aapa, american2018ashg, board1998aaa, fuentes2019aapa, nhgri2023, rotimi2022ashg, wynshawboris2020ashg}, undermining the core assumption of race science. As described by Winston, however, "[s]ince the 1990s, scientific racism has evolved in ways that sidestep or blunt the charge of essentialist categorization" \cite[p. 427]{WinstonWhyMainstreamResearch2020}, such as shifting the rhetoric away from race to concepts such as "national IQ" that are deployed to preserve an essentialist categorization of different human populations. Professional psychological organizations routinely publish statements rejecting racism or acknowledging psychology's complicity in systemic racism \cite{american2021apology, aps2021statement}, but stop short of rejecting race as a meaningful biological categorization of people, allowing for race science research to be smuggled into psychology and giving the appearance of continued scientific legitimacy.
